Concurrent Programming in Harmony

### 2021-02-23 book/paper.tex
Latest

@@

\section*{Exercises}
\begin{problems}
\item Run \autoref{fig:cs} using Harmony.  As there is no protection of the critical
section, mutual exclusion is violated,
the assertion should fail, and a trace should be reported.
Now insert
\begin{code}
\texttt{await False}
\end{code}
just before entering the critical section
in \autoref{fig:cs} and run Harmony again.
Mutual exclusion is guaranteed but progress is violated.
Harmony should print a trace
to a state from which a terminating state cannot be reached.
Describe in English the difference in the failure reports before
and after inserting the code.
\item See if you can come up with some different approaches that satisfy both
mutual exclusion and progress.  Try them with Harmony and see if they work or not.
If they don't, try to understand why.
Do not despair if you can't figure out how to develop a solution that satisfies both
mutual exclusion and progress---as
we will find out, it is possible but not obvious.
\end{problems}

@@

\index{data race}
When there is a bug in a program because some code omitted obtaining
a lock before accessing a shared data structure, that is known as a
\emph{data race}.
More precisely, a data race happens when there is a state in which multiple
threads are trying to access the same variable, and at least one of those
accesses updates the variable.
In many environments, including C and Java programs, the behavior of
concurrent load and store operations have tricky or even undefined
semantics.  One should therefore avoid data races, which is why Harmony
reports them even though Harmony has sequentially consistent memory.

>>>>

\newglossaryentry{data race}
  name=data race,
  description={is when there are two or more threads concurrently accessing a
    shared variable, at least one of which is an update to the variable}

@@

\section*{Exercises}
\begin{problems}
\item Run \autoref{fig:cs} using Harmony.  As there is no protection of the critical
section, mutual exclusion is violated,
the assertion should fail, and a trace should be reported.
Now insert
\begin{code}
\texttt{await False}
\end{code}
just before entering the critical section
in \autoref{fig:cs} and run Harmony again.
Mutual exclusion is guaranteed but progress is violated.
Harmony should print a trace
to a state from which a terminating state cannot be reached.
Describe in English the difference in the failure reports before
and after inserting the code.
\item See if you can come up with some different approaches that satisfy both
mutual exclusion and progress.  Try them with Harmony and see if they work or not.
If they don't, try to understand why.
If you get \emph{active busy waiting} or \emph{data race} reports, you probably
found a correct solution; you'll learn later how to suppress those.
Do not despair if you can't figure out how to develop a solution that satisfies both
mutual exclusion and progress---as we will find out, it is possible but not obvious.
\end{problems}

@@

\index{data race}%
\glsadd{data race}%
When there is a bug in a program because some code omitted obtaining
a lock before accessing a shared data structure, that is known as a
\emph{data race}.
More precisely, a data race happens when there is a state in which multiple
threads are trying to access the same variable, and at least one of those
accesses updates the variable.
In many environments, including C and Java programs, the behavior of
concurrent load and store operations have tricky or even undefined
semantics.  One should therefore avoid data races, which is why Harmony
reports them even though Harmony has sequentially consistent memory.
